---
output:
  word_document: default
  html_document: default
  pdf_document: default
editor_options:
  markdown:
    wrap: 72
---

```{r}
library(quantmod)
library(tidyverse)
library(lubridate)
library(xts)
library(zoo)
library(ggplot2)
library(sandwich)
library(lmtest)


```

#10-yr Treasury Yield [not seasonally adjusted (%)], CPI Index, Crude
Oil Prices [not seasonally adjusted (\$ per barrel)] #2010-Present

```{r}
#download data
getSymbols(
  Symbols = c("DGS10","CPIAUCSL","DCOILWTICO"), src="FRED", 
  auto.assign = TRUE, 
  from = as.Date("2012-01-01")
  )

#need to fix missing dates and change oil and treasury to average monthly?
dgs10 = apply.monthly(DGS10, mean, na.rm=TRUE) # in %
cpi = CPIAUCSL #already monthly
oil = apply.monthly(DCOILWTICO, mean, na.rm=TRUE) #in $/barrel

index(cpi) = ceiling_date(index(cpi),"month") - days(1)
index(dgs10) = ceiling_date(index(dgs10), "month") - days(1)
index(oil) = ceiling_date(index(oil), "month") - days(1)

inflation_mom = diff(log(cpi)) *100
colnames(inflation_mom) = "inflation_mom"

#final pivot
# macro = merge(DGS10, DCOILWTICO, CPIAUCSL, join = "inner")
# 
# macro_df = macro %>%
#   data.frame(date=index(macro)) %>%
#   as_tibble()
# 
# macro_long = macro_df %>%
#   select(date, rate10 = DGS10, oil_price = DCOILWTICO, inflation) %>%
#   pivot_longer(-date, names_to = "variable", values_to = "value")
# 
# ggplot(macro_long, aes(x=date, y=value, color=variable)) +
#   geom_line(size=0.8) +
#   facet_wrap(~variable, scales="free_y", ncol=1) +
#   labs(title="Macro Trends: Interest Rates, Oil Prices, Inflation",
#        x="Date", y="Value") +
#   theme_minimal() +
#   theme(legend.position="none")

##merging macro data
macro = merge(dgs10, oil, inflation_mom, join = "inner") %>%
  setNames(c("treasury_10yr", "oil_price", "inflation_mom"))
```

#Move on to ETFs Info (Renewables vs. Brown Energy)

```{r}
brown_tickers = c("XLE", "VDE")
clean_tickers = c("ICLN", "TAN", "QCLN")


getSymbols(c(clean_tickers, brown_tickers),
           src="yahoo", from="2012-01-01", auto.assign = TRUE)

to_mon_ret = function(ticker) {
  x = get(ticker)
  periodReturn(Ad(x), period = "monthly", type = "log")   # log returns play nicer in regressions
}

clean_ret_mat = do.call(merge, lapply(clean_tickers, to_mon_ret))

clean_ret = xts(rowMeans(clean_ret_mat, na.rm = TRUE), order.by = index(clean_ret_mat))
colnames(clean_ret) = "clean_returns"

brown_ret_mat = do.call(merge, lapply(brown_tickers, to_mon_ret))

brown_ret = xts(rowMeans(brown_ret_mat, na.rm = TRUE), order.by = index(brown_ret_mat))
colnames(brown_ret) = "brown_returns"
##month end alignment 
index(clean_ret) = ceiling_date(index(clean_ret), "month") - days(1)
index(brown_ret) = ceiling_date(index(brown_ret), "month") - days(1)

relative_ret = clean_ret - brown_ret
colnames(relative_ret) = "relative_returns"


```

#Consolidate macro data and ETF returns

```{r}
macro$relative_returns = relative_ret
macro = na.omit(macro)
head(macro)
```

#transforming into data frame

```{r}
macro_df = data.frame(date = as.Date(index(macro)),
              treasury10 = as.numeric(macro$treasury_10yr),
              oil_price = as.numeric(macro$oil_price),
              inflation = as.numeric(macro$inflation_mom),
              relative_returns = as.numeric(macro$relative_returns)
)

macro_df$clean_returns = as.numeric(clean_ret[index(macro)])
macro_df$brown_returns = as.numeric(brown_ret[index(macro)])

head(macro_df)


```

#stat summary basic

```{r}
mean_rel = mean(macro_df$relative_returns)
sd_rel = sd(macro_df$relative_returns)
sharpe_rel = mean_rel / sd_rel * sqrt(12)

cat("Start:", min(macro_df$date), "\n")

cat("End:", max(macro_df$date), "\n")

cat("Number of months:", nrow(macro_df), "\n") #188 months

cat("Mean monthly relative return:", mean_rel, "\n") #on average clean energy overperforms brown energy by 0.17% per month ------ (1-0.001699274 )^12 - 1 =-0.02058295 around 2.06% / yr from 2012-2025
cat("Volatility (SD):", sd_rel, "\n") #relative performance fluctuates very wildly month to month (9.25% monthly volatility)
cat("Annualized Sharpe (Clean - Brown):", sharpe_rel, "\n") 



```

```{r}

x1 = macro_df$oil_price
x2 = macro_df$inflation
x3 = macro_df$treasury10
y = macro_df$relative_returns


# Pearson
cat("--- Pearson correlations with relative_returns ---\n")
cat("Oil price:", cor(y, x1, use = "complete.obs", method = "pearson"), "\n") #\n creates new line
cat("Inflation:", cor(y, x2, use = "complete.obs", method = "pearson"), "\n")
cat("10Y rate:", cor(y, x3, use = "complete.obs", method = "pearson"), "\n")


# Spearman
cat("--- Spearman correlations with relative_returns ---\n")
cat("Oil price:", cor(y, x1, use = "complete.obs", method = "spearman"), "\n")
cat("Inflation:", cor(y, x2, use = "complete.obs", method = "spearman"), "\n")
cat("10Y rate:", cor(y, x3, use = "complete.obs", method = "spearman"), "\n")


#oil prices increase, brown energy rally bc higher profit margins? for fossil fuel stakeholders
#inflation increase, Renewables are CAPITAL-INTENSIVE (project finance + governemnt incentives often increases cost of borrowing and costs)
#rising rate hurt growth stock valuations (long duration growth assets)
  #fossil fuel companies have short duration and immediate cash flow

```

#regime analysis #idea is to split sample into two groups high and low
oil months where they are either below or above median #then we would
compare mean relative returns across groups

```{r}
#define regimes using median
#conditional expectation (E[return | regime])
oil_median = median(macro_df$oil_price, na.rm = TRUE)
macro_df$oil_regime = ifelse(macro_df$oil_price > oil_median, "High","Low")

#table(macro_df$oil_regime)

#summary
mean_by_regime = tapply(macro_df$relative_returns,
                        macro_df$oil_regime,
                        mean,
                        na.rm=TRUE
                        )
sd_by_regime = tapply(macro_df$relative_returns,
                          macro_df$oil_regime,
                        sd,
                        na.rm=TRUE
                        )
sharpe_by_regime = mean_by_regime / sd_by_regime * sqrt(12)
cat("--- High vs Low Oil pricing Regime stats ---\n")
cat("mean: ", mean_by_regime,"\n")
cat("sd:    ", sd_by_regime,"\n")
cat("sharpe:", sharpe_by_regime, "\n")

#green energy underperforms during high oil prices by 1.63
#Green performs well during low by only 2.07

#both regimes are volatile, HOWEVER high oil periods are more CHAOTIC
#high oil price periods include supply shocks, geopolitical stress, OPEC policy/regulation...


#


```

#Boxplot of Relative Returns by REGIME

```{r}
boxplot(relative_returns ~ oil_regime,
        data = macro_df,
        main = "Relative returns by Oil Regime",
        xlab = "Oil Regime",
        ylab = "Monthly Relative Return",
        col = c("lightblue", "lightpink"))
abline(h=0, col = "grey")
        

```

#cumulative performance plots (clean vs brown vs relative)

```{r}
#basically cumulative return from monthly returns

cumulative_from_simple = function(r) {
  r_no_na = ifelse(is.na(r), 0, r)
  cumprod(1 + r_no_na) - 1
}

clean_cumulative = cumulative_from_simple(macro_df$clean_returns)

brown_cumulative = cumulative_from_simple(macro_df$brown_returns)

rel_cumulative = cumulative_from_simple(macro_df$relative_returns)

plot(macro_df$date, clean_cumulative,
     type = "l", col = "blue", lwd = 1.5,
     xlab = "Date", ylab = "Cumulative Return",
     main = "Cumulative Performance: Clean v Brown")
lines(macro_df$date, brown_cumulative, col = "red", lwd = 1.5)
legend("bottomleft",
       legend = c("Clean", "Brown"),
       col    = c("blue", "red"),
       lty    = 5, bty = "n")
abline(h = 0, col = "grey")

plot(macro_df$date, rel_cumulative,
     type = "l", col = "darkgreen", lwd = 2,
     xlab = "Date", ylab = "Cumulative return",
     main = "Cumulative performance: Clean - Brown")
abline(h = 0, col = "grey")

```

#Rolling 12-mo Sharpe Ratio of Relative Returns

```{r}
#concepts: rolling window = compute stats over Last N months
#Sharpe ratio = mean/SD in that window * sqrt12

rolling_mean_12 = rollapply(macro_df$relative_returns,
                            width = 12, FUN = mean,
                            by = 1, align = "right",
                            fill = NA, na.rm = TRUE
                            )
rolling_sd_12 = rollapply(macro_df$relative_returns,
                            width = 12, FUN = sd,
                            by = 1, align = "right",
                            fill = NA, na.rm = TRUE
                            )
rolling_sharpe_12 = rolling_mean_12 / rolling_sd_12 * sqrt(12)

plot(macro_df$date, rolling_sharpe_12,
     type = "l", lwd = 2, col = "darkgreen",
     xlab = "Date", ylab = "Rolling Sharpe",
     main = "Rolling 12-Month Sharpe (Clean - Brown)"
     )
abline(h=0, col="black", lwd = 2)


```

###----Simple Linear Regression + Adding Momentum \##

```{r}
#Using past macro data to explain current returns...

#Adding 3m and 6m momentum factors earlier for easier 
relative_momentum_3m = rollapply(macro_df$relative_returns,
                            width = 3, FUN = mean,
                            by = 1, align = "right",
                            fill = NA, na.rm = TRUE
                            )
relative_momentum_6m = rollapply(macro_df$relative_returns,
                            width = 6, FUN = mean,
                            by = 1, align = "right",
                            fill = NA, na.rm = TRUE
                            )



lag1 = function(x) c(NA, x[-length(x)])



regression_df = data.frame(
  date = macro_df$date,
  relative_returns = macro_df$relative_returns,
  oil_price_lag1 = lag1(macro_df$oil_price),
  inflation_lag1 = lag1(macro_df$inflation),
  treasury10_lag1 = lag1(macro_df$treasury10),
  relative_momentum_3m = lag1(relative_momentum_3m),
  relative_momentum_6m = lag1(relative_momentum_6m)
) %>% 
  na.omit()

head(regression_df)


```

#Basic Regression w/ lag 1 base macro factors

```{r}
model1 = lm(relative_returns ~ oil_price_lag1 + inflation_lag1 + treasury10_lag1, data = regression_df)

summary(model1)
```

#Regression with base macro + 3 and 6 mo relative momentum

```{r}
#rolling means


model2 = lm(relative_returns ~ oil_price_lag1 + inflation_lag1 + treasury10_lag1 +relative_momentum_3m + relative_momentum_6m,
            data = regression_df)

summary(model2)
#coefficients remain consistent with economic intuition
#positive momentum coefficients suggests that recent relative outperformance tends to persist

#though, none f the predictors are statistically significant.

#Can suggest that clean-brown performance are influenced by macro effects thru regime shifts than month to month linear predictability.

```

#Next to make inference more robust? HAC/Newey West standard errors #NW
adjusts standard errors so p-values are more reliable

#testing for heteroskedasticity and autocorrelation

```{r}
#sanity check
dwtest(model2)
bgtest(model2, order=6)
bptest(model2)


#cant reject null for each test, no need for newey west adjustments
```

#classification #datas setup

```{r}
#building classification dataset
regression_df = regression_df[order(regression_df$date),]

#create binary target for relative return of next month
target_next = c(regression_df$relative_returns[-1] > 0, NA)

classification_df = regression_df
classification_df$target_next = target_next
classification_df = classification_df[!is.na(classification_df$target_next), ] #drop last row (there is no next month target)


```

#logistic regression (walk-forward)

```{r}
feature_cols = c("oil_price_lag1",
                  "inflation_lag1",
                  "treasury10_lag1",
                  "relative_momentum_3m",
                  "relative_momentum_6m")

#column of logit probabilities

n= nrow(classification_df)

logit_prob = rep(NA_real_, n)

start_oos = 36 

#walk forward loop

for (i in start_oos:(n-1)) {
  train_rows = 1:i
  test_row = i + 1
  
  train_data =  classification_df[train_rows, c("target_next", feature_cols)] #train rows are based on target and features
  test_data = classification_df[test_row, feature_cols, drop=FALSE] #predict based on features only
  
  logistic_model = glm( target_next ~., data = train_data, family = binomial()
                        ) #fit the model
  
  logit_prob[test_row] = predict(logistic_model, newdata = test_data, type = "response") #predict prob next month
   
  
}

classification_df$logit_prob = logit_prob

#check classification accuracy
oos_index = (start_oos + 1):n
prediction_label = logit_prob[oos_index] > 0.5
actual_label = classification_df$target_next[oos_index]
logistic_accuracy = mean(prediction_label == actual_label, na.rm=TRUE)

logistic_accuracy
#0.4710744


```

#regime, momentum, probability signals

```{r}
#regime signal setup
classification_df$oil_price = macro_df$oil_price[match(classification_df$date, macro_df$date)]

macro_signal = ifelse(classification_df$oil_price <= oil_median, 1, 0)

#momentum setup
momentum3_signal = ifelse(classification_df$relative_momentum_3m > 0, 1, 0)
momentum6_signal = ifelse(classification_df$relative_momentum_6m > 0, 1, 0)

#logistic probability signal
#set up confidence of >=0.50

log_signal = ifelse(classification_df$logit_prob >= 0.5, 1, 0)

classification_df$macro_signal = macro_signal
classification_df$momentum3_signal = momentum3_signal
classification_df$momentum6_signal = momentum6_signal
classification_df$log_signal = log_signal

#compile signals into score container
classification_df$signal_score = classification_df$macro_signal + classification_df$momentum3_signal + classification_df$momentum6_signal +
  classification_df$log_signal

classification_df$long_clean = ifelse(classification_df$signal_score >= 3, 1, 0)


table(classification_df$signal_score)
table(classification_df$long_clean)

```

#strategy test

```{r}
#compare benchmark and our strategy
#strategy is to long only when long_clean = 1

benchmark = classification_df$relative_returns

strategy = classification_df$relative_returns * classification_df$long_clean


#we are going to use the cumulative_from_simple function

strategy_cumulative = cumulative_from_simple(strategy)

benchmark_cumulative = cumulative_from_simple(benchmark)

plot(classification_df$date, benchmark_cumulative,
     type = "l", lwd = 2, col = "blue",
     xlab ="Date", ylab = "Cumulative Return",
     main = "Cumulative Performance: Benchmark Always Long vs Long when Signal")
lines(classification_df$date, strategy_cumulative,
      col="black", lwd = 2)
abline(h=0,col="grey")
legend("topleft",
       legend = c("Always Long","Signal Based Long"), col = c("blue","black"), lty = 1,bty ="n")

#With strategy that uses atleast 3 score signals

```

#Sharpe Ratios

```{r}

strategy_mean = mean(strategy, na.rm=TRUE)
strategy_sd = sd(strategy, na.rm=TRUE)
strategy_sharpe = strategy_mean / strategy_sd *sqrt(12)

benchmark_mean = mean(benchmark, na.rm=TRUE)
benchmark_sd = sd(benchmark, na.rm=TRUE)
benchmark_sharpe = benchmark_mean / benchmark_sd *sqrt(12)


cat("Strategy Sharpe:", strategy_sharpe, "\n")
cat("Benchmark Sharpe:", benchmark_sharpe,"\n")
#strategy sharpe > benchmark sharpe

```

###further improvements? #adding additional possibly useful features

```{r}
#adding VIX (market volatility indicator; uncertainty may hurt clean performance)

getSymbols("^VIX", src = "yahoo", from = "2012-01-01", auto.assign = TRUE)

vix_daily = Ad(VIX)
vix = apply.monthly(vix_daily, mean, na.rm = TRUE)

index(vix) = ceiling_date(index(vix), "month") - days(1)
colnames(vix) = "vix"
head(vix)

macro = merge(macro, vix, join = "inner")
macro_df$vix = as.numeric(vix[index(macro)])

vix_lag_full = lag1(macro_df$vix)

regression_df$vix_lag1 = vix_lag_full[match(regression_df$date, macro_df$date)]



# head(regression_df)
#adding spy to find out whether spy can favor clean etfs (if market favors growht stocks)
getSymbols("SPY", src = "yahoo", from = "2012-01-01", auto.assign = TRUE)

spy_return = periodReturn(Ad(SPY), period = "monthly", type = "log")

index(spy_return) = ceiling_date(index(spy_return), "month") -days(1)

macro_df$clean_spy = macro_df$clean_returns - as.numeric(spy_return[index(macro_df$date)])


clean_spy_full_lag =  lag1(macro_df$clean_spy)


regression_df$clean_spy_lag1 = clean_spy_full_lag[match(regression_df$date, macro_df$date)]


#put into classificaiton table
classification_df$vix_lag1 = regression_df$vix_lag1[match(classification_df$date, regression_df$date)]

classification_df$clean_spy_lag1 = regression_df$clean_spy_lag1[match(classification_df$date, regression_df$date)]

```

#rerun logisitic regression model

```{r}
feature_cols = c(
  "oil_price_lag1",
  "inflation_lag1",
  "treasury10_lag1",
  "relative_momentum_3m",
  "relative_momentum_6m",
  "vix_lag1",
  "clean_spy_lag1"
)

n= nrow(classification_df)

logit_prob = rep(NA_real_, n)

start_oos = 36 

#walk forward loop

for (i in start_oos:(n-1)) {
  train_rows = 1:i
  test_row = i + 1
  
  train_data =  classification_df[train_rows, c("target_next", feature_cols)] #train rows are based on target and features
  test_data = classification_df[test_row, feature_cols, drop=FALSE] #predict based on features only
  
  logistic_model = glm(target_next ~., data = train_data, family = binomial()
                        ) #fit the model
  
  logit_prob[test_row] = predict(logistic_model, newdata = test_data, type = "response") #predict prob next month
   
  
}

classification_df$logit_prob = logit_prob

oos_index = (start_oos + 1):n
prediction_label = logit_prob[oos_index] > 0.5
actual_label = classification_df$target_next[oos_index]
logistic_accuracy2 = mean(prediction_label == actual_label, na.rm=TRUE)

logistic_accuracy
logistic_accuracy2
#close to same figure prior to adding features
```

#re-evaluate strategy

```{r}
log_signal = ifelse(classification_df$logit_prob >= 0.5, 1, 0)

classification_df$signal_score = macro_signal + momentum3_signal + momentum6_signal + log_signal
#consider loosening score variable to 2
classification_df$long_clean = ifelse(classification_df$signal_score >= 3, 1, 0)

strategy2  = classification_df$relative_returns * classification_df$long_clean

strategy2_cumulative  = cumulative_from_simple(strategy2)
benchmark_cumulative = cumulative_from_simple(benchmark)

plot(classification_df$date, benchmark_cumulative,
     type = "l", lwd = 2, col = "blue",
     xlab ="Date", ylab = "Cumulative Return",
     main = "Cumulative Performance: Benchmark Always Long vs Long when Signal (New)")
lines(classification_df$date, strategy2_cumulative,
      col="black", lwd = 2)
abline(h=0,col="grey")
legend("topleft",
       legend = c("Always Long","Signal Based Long"),
       col = c("blue","black"), lty = 1,bty ="n")
strategy2_mean   = mean(strategy2,   na.rm = TRUE)
strategy2_sd     = sd(strategy2,     na.rm = TRUE)
strategy2_sharpe = strategy2_mean / strategy2_sd * sqrt(12)

cat("Strategy Sharpe:", strategy_sharpe, "\n")
cat("Strategy 2 Sharpe:", strategy2_sharpe, "\n")
cat("Benchmark Sharpe:", benchmark_sharpe, "\n")

#accuracy barely improved, sharpe ratio worsened after feature addition 
```

#try utilizing xgboost into our model

```{r}
library(xgboost)

X = as.matrix(classification_df[, feature_cols])
y = as.numeric(classification_df$target_next)  # 1/0
n = nrow(X)

xgb_prob = rep(NA_real_, n)

for (i in start_oos:(n - 1)) {
  train_index = 1:i
  test_index  = i + 1
  
  dtrain = xgb.DMatrix(data = X[train_index,], label = y[train_index])
  
  dtest = xgb.DMatrix(data = X[test_index, , drop=FALSE])
  
  params = list(
    objective = "binary:logistic",
    eta = 0.1,
    max_depth = 3,
    subsample = 0.8,
    colsample_bytree = 0.8,
    eval_metric = "logloss"
  )
  
  xgb_model = xgb.train(
    params = params,
    data = dtrain,
    nrounds = 100,
    verbose = 0
  )
  xgb_prob[test_index] = predict(xgb_model, newdata = dtest)
}

classification_df$xgb_prob = xgb_prob

xgb_prediction = classification_df$xgb_prob[oos_index] > 0.5
actual = classification_df$target_next[oos_index]

xgb_accuracy = mean(xgb_prediction == actual, na.rm = TRUE)
logistic_accuracy
logistic_accuracy2
xgb_accuracy


#### testing xgb probabilities

xgb_signal = ifelse(classification_df$xgb_prob >= 0.5, 1, 0)

classification_df$signal_score = macro_signal + momentum3_signal + momentum6_signal + xgb_signal
#consider loosening score variable to 2
classification_df$long_clean = ifelse(classification_df$signal_score >= 3, 1, 0)

strategy3  = classification_df$relative_returns * classification_df$long_clean

strategy3_cumulative  = cumulative_from_simple(strategy3)
# benchmark_cumulative = cumulative_from_simple(benchmark)

plot(classification_df$date, benchmark_cumulative,
     type = "l", lwd = 2, col = "blue",
     xlab ="Date", ylab = "Cumulative Return",
     main = "Cumulative Performance: Benchmark Always Long vs Long when Signal (New)")
lines(classification_df$date, strategy3_cumulative,
      col="black", lwd = 2)
abline(h=0,col="grey")
legend("topleft",
       legend = c("Always Long","Signal Based Long"),
       col = c("blue","black"), lty = 1,bty ="n")
strategy3_mean   = mean(strategy3,   na.rm = TRUE)
strategy3_sd     = sd(strategy3,     na.rm = TRUE)
strategy3_sharpe = strategy3_mean / strategy3_sd * sqrt(12)

cat("Strategy Sharpe:", strategy_sharpe, "\n")
cat("Strategy 2 Sharpe:", strategy2_sharpe, "\n")
cat("Strategy 3 Sharpe:", strategy3_sharpe, "\n")
cat("Benchmark Sharpe:", benchmark_sharpe, "\n")
#
```
#finding best combination of probability threshold and signal score

```{r}

thresholds = seq(0.40, 0.70, by = 0.01)   # prob cutoff for logit signal
score_cutoffs = 1:4                       # how many signals needed to go long

results = expand.grid(
  threshold = thresholds,
  score_cutoff = score_cutoffs)

results$sharpe = NA_real_

results$avg_exposure = NA_real_   # fraction of months in the market


for (i in seq_len(nrow(results))) {

  th = results$threshold[i]
  cut = results$score_cutoff[i]

  # logistic probability signal at this threshold
  xgb_signal = ifelse(classification_df$xgb_prob >= th, 1, 0)

  # recompute score
  signal_score = macro_signal +
                 momentum3_signal +
                 momentum6_signal +
                 xgb_signal

  long_clean = ifelse(signal_score >= cut, 1, 0)

  # strategy returns
  strategy_ret = classification_df$relative_returns * long_clean

  # if sd is zero (e.g. no trades), Sharpe = NA
  if (sd(strategy_ret, na.rm = TRUE) > 0) {
    sharpe = mean(strategy_ret, na.rm = TRUE) / sd(strategy_ret, na.rm = TRUE) * sqrt(12)
  } else {
    sharpe = NA_real_
  }

  results$sharpe[i] = sharpe
  results$avg_exposure[i] = mean(long_clean, na.rm = TRUE)   # % of months in the market
}

# find best combo by Sharpe
best_row_index = which.max(results$sharpe)
best_combo = results[best_row_index, ]

best_combo
```

#looking at best combination results

```{r}

best_threshold   = best_combo$threshold
best_score_cut   = best_combo$score_cutoff


xgb_signal_best = ifelse(classification_df$xgb_prob >= best_threshold, 1, 0)

classification_df$signal_score_best =
  macro_signal +
  momentum3_signal +
  momentum6_signal +
  xgb_signal_best

classification_df$long_clean_best =
  ifelse(classification_df$signal_score_best >= best_score_cut, 1, 0)

# strategy returns with best rule
strategy_best = classification_df$relative_returns * classification_df$long_clean_best

strategy_best_mean   = mean(strategy_best, na.rm = TRUE)
strategy_best_sd     = sd(strategy_best,   na.rm = TRUE)
strategy_best_sharpe = strategy_best_mean / strategy_best_sd * sqrt(12)

cat("Best threshold:", best_threshold, "\n")
cat("Best score cutoff:", best_score_cut, "\n")
cat("Best strategy Sharpe:", strategy_best_sharpe, "\n")
cat("Benchmark Sharpe:", benchmark_sharpe, "\n")   # assuming benchmark_sharpe from before

# cumulative performance plot
strategy_best_cumulative  = cumulative_from_simple(strategy_best)
# we already have benchmark_cumulative

plot(classification_df$date, benchmark_cumulative,
     type = "l", lwd = 2, col = "blue",
     xlab = "Date", ylab = "Cumulative Return",
     main = "Benchmark vs Optimized Signal-Based Strategy")
lines(classification_df$date, strategy_best_cumulative,
      col = "black", lwd = 2)
abline(h = 0, col = "grey")
legend("topleft",
       legend = c("Always Long", "Signal-Based (Optimized)"),
       col    = c("blue", "black"),
       lty    = 1, bty   = "n")
```

#what does combining both log_prob and xgb_prob as signals maybe as averages
```{r}
classification_df$ensemble_prob = rowMeans(
  cbind(classification_df$logit_prob, classification_df$xgb_prob),
  na.rm = TRUE
)

ensemble_threshold = 0.50  

classification_df$ensemble_signal = ifelse(
  classification_df$ensemble_prob >= ensemble_threshold, 1, 0
)

classification_df$signal_score_ensemble =
  macro_signal +
  momentum3_signal +
  momentum6_signal +
  classification_df$ensemble_signal

# Choose a score cutoff – start with 3 like before
score_cutoff_ensemble = 3

classification_df$long_clean_ensemble = ifelse(
  classification_df$signal_score_ensemble >= score_cutoff_ensemble, 1, 0
)

strategy_ret_ensemble = classification_df$relative_returns * classification_df$long_clean_ensemble


strategy_cumulative_ensemble = cumulative_from_simple(strategy_ret_ensemble)


# Sharpe
strategy_mean_ensemble   <- mean(strategy_ret_ensemble, na.rm = TRUE)
strategy_sd_ensemble     <- sd(strategy_ret_ensemble,   na.rm = TRUE)
strategy_sharpe_ensemble <- strategy_mean_ensemble / strategy_sd_ensemble * sqrt(12)

plot(classification_df$date, benchmark_cumulative,
     type = "l", lwd = 2, col = "blue",
     xlab = "Date", ylab = "Cumulative Return",
     main = "Benchmark vs Optimized Signal-Based Strategy")
lines(classification_df$date, strategy_cumulative_ensemble,
      col = "black", lwd = 2)
abline(h = 0, col = "grey")
legend("topleft",
       legend = c("Always Long", "Signal-Based (Ensemble)"),
       col    = c("blue", "black"),
       lty    = 1, bty   = "n")


strategy_sharpe
strategy2_sharpe
strategy3_sharpe
strategy_sharpe_ensemble
benchmark_sharpe

#highest sharpe!
```



#output to csv for tableau use and learning
```{r}
# build returns & cumulative series
benchmark_ret = classification_df$relative_returns
strategy_ret_ensemble = classification_df$relative_returns * classification_df$long_clean_ensemble

classification_df$xgb_signal = xgb_signal
benchmark_cum = cumulative_from_simple(benchmark_ret)
ensemble_cum  = cumulative_from_simple(strategy_ret_ensemble)

# attach to classification_df with consistent column names
classification_df <- classification_df %>%
  mutate(
    benchmark_ret        = benchmark_ret,
    strategy_ret_ensemble = strategy_ret_ensemble,
    benchmark_cumulative  = benchmark_cum,
    ensemble_cumulative  = ensemble_cum)

tableau_df <- macro_df %>%
  # keep macro + clean/brown/relative returns
  select(
    date,
    treasury10,
    oil_price,
    inflation,
    clean_returns,
    brown_returns,
    relative_returns,
    vix,
    clean_spy
  ) %>%
  # join modeling/signal info by date
  left_join(
    classification_df %>%
      select(all_of(c(
        "date",
        "target_next",
        "logit_prob",
        "xgb_prob",
        "ensemble_prob",
        "macro_signal",
        "momentum3_signal",
        "momentum6_signal",
        "log_signal",           # logistic binary column
        "xgb_signal",           # xgb binary column
        "ensemble_signal",
        "signal_score_ensemble",
        "long_clean_ensemble",
        "benchmark_ret",
        "strategy_ret_ensemble",
        "benchmark_cumulative",
        "ensemble_cumulative"
      ))),
    by = "date"
  )

write.csv(
  tableau_df,
  "tableau_clean_vs_brown_framework.csv",
  row.names = FALSE
)

```




